---
title: "factor_analysis"
output: pdf_document
date: "2025-01-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#svd analysis

## get the data
``` {r, message = FALSE}

library(readxl)
library(statnet)
library(igraph)
library(intergraph)

data_usa <- read_excel("/Users/wumengmeng/Desktop/project-MediaAgenda/two-mode/step1testcode/step2_code_data/exploratory analysis/ce_cor_matrix_usa0104.xlsx")
matrix_two_mode <- as.matrix(data_usa)

tdata_usa <- read_excel("/Users/wumengmeng/Desktop/project-MediaAgenda/two-mode/step1testcode/step2_code_data/exploratory analysis/transposed_ce_matrix_usa0105.xlsx")
matrix_two_mode <- as.matrix(data_usa)
tmatrix_two_mode <- as.matrix(tdata_usa)
```


```{r}
#construct a function to convert data type
convert_to_numeric_matrix <- function(matrix_two_mode) {
 
  
  issue_labels <- colnames(matrix_two_mode)[-1]
  
  country_labels <- matrix_two_mode[, 1]
  
  numeric_data <- matrix_two_mode[, -1]
  numeric_data <- apply(numeric_data, c(1, 2), trimws)
  
  numeric_matrix <- apply(numeric_data, 2, as.numeric)
  length(country_labels) 
  dim(numeric_matrix)
  rownames(numeric_matrix) <- country_labels
  colnames(numeric_matrix) <- issue_labels
  return(numeric_matrix)
}

```


```{r}
#convert data type in matrix

numeric_matrix <- convert_to_numeric_matrix(matrix_two_mode)

tnumeric_matrix <- convert_to_numeric_matrix(tmatrix_two_mode)
print(tnumeric_matrix)
```

```{r}


# 步骤 1: 标准化矩阵 residuals
O <- numeric_matrix / sum(numeric_matrix)         # 原始概率矩阵
E <- rowSums(O) %o% colSums(O)                   # 期望值矩阵
Z <- (O - E) / sqrt(E)                           # 标准化残差矩阵

# 步骤 2: 计算SVD
SVD <- svd(Z)

# 输出 SVD 结果
cat("Singular values (d):\n")
print(SVD$d)   # 奇异值
cat("\nLeft singular vectors (u):\n")
print(SVD$u)   # 左奇异向量
cat("\nRight singular vectors (v):\n")
print(SVD$v)   # 右奇异向量


# 步骤 4: 计算方差解释比例
variance_explained <- prop.table(SVD$d^2)  # 方差比例
cat("\nVariance explained by each singular value:\n")
print(variance_explained)

# 步骤 5: 可选 - 只使用前 k 个奇异值进行降维重构
k <- 4  # 设置保留的维度数
U_k <- SVD$u[, 1:k]  # 左奇异向量 (行向量)
V_k <- SVD$v[, 1:k]  # 右奇异向量 (列向量)
Sigma_k <- diag(SVD$d[1:k])  # 前两个奇异值对角矩阵

# 步骤 2: 计算 Row Scores 和 Column Scores
# Row Scores = U_k * Sigma_k
row_scores <- U_k %*% Sigma_k

# Column Scores = V_k * Sigma_k
column_scores <- V_k %*% Sigma_k

# 输出结果
cat("\nRow Scores (First 2 dimensions):\n")
print(row_scores)

cat("\nColumn Scores (First 2 dimensions):\n")
print(column_scores)
```
```{r}
#convert data type in matrix

numeric_matrix <- convert_to_numeric_matrix(matrix_two_mode)
```



```{r}
# 输出结果
cat("\nRow Scores (First 4 dimensions):\n")
print(row_scores)

cat("\nColumn Scores (First 4 dimensions):\n")
print(column_scores)

write.csv(row_scores, file = "row_scores.csv", row.names = rownames(numeric_matrix))
write.csv(column_scores, file = "column_scores.csv", row.names = colnames(numeric_matrix))
```

```{r}
# Clustering function
cluster_data <- function(scores) {
  # Initialize groups as a list with four empty groups
  groups <- list(
    group1 = integer(0),
    group2 = integer(0),
    group3 = integer(0),
    group4 = integer(0)
  )
  
  # Iterate through each row of the scores matrix
  for (i in 1:nrow(scores)) {
    max_dim <- which.max(scores[i, ]) # Find the index of the max value (1-based in R)
    group_name <- paste0("group", max_dim) # Determine the group name
    groups[[group_name]] <- c(groups[[group_name]], i) # Append the row index to the group
  }
  
  return(groups) # Return the groups
}
row_groups = cluster_data(row_scores)
col_groups = cluster_data(column_scores)
row_groups
col_groups


  
```
```{r}

print(rownames(numeric_matrix))
```

#factor analysis



```{r}

# 执行因子分析，设定因子数 (k = 4)
result_f <- factanal(numeric_matrix, factors = 4, rotation = "varimax")

# 查看因子载荷矩阵（loadings）
print("Factor loadings:")
print(result_f$loadings)

# 查看因子解释的方差比例
print("Proportion of variance explained:")
print(result_f$variance)



```
Factor1: 社会服务与基础设施 （合并了svd的group1&4）
话题：
healthcare (载荷：0.746)
publicorder (载荷：0.727)
civil_liberties (载荷：0.757)
education (载荷：0.744)
infrastructure (载荷：0.753)
poverty (载荷：0.724)
disaster (载荷：0.717)

Factor2: 经济与环境（合并了svd的group2&3）
话题：
economy (载荷：0.768)
military&terrorism (载荷：0.746)
borderissues (载荷：0.749)
environment (载荷：0.730)
politics (载荷：0.705)
religion (载荷：0.708)
media/internet (载荷：0.711)


#对于转置后的矩阵，协方差矩阵是奇异的（不可逆），意味着协方差矩阵没有满秩，其数学特性导致它无法计算逆矩阵


